Assessing the Impact of MDE and Code Generation on the Sustainability of Scientific Computing Software: A Research Proposal

### 2019-09-18 ResearchProposal.tex
Edits to research proposal

DSLs and code/document generation provide a transformative technology for
documentation, design and verification \citep{JohansonAndHasselbring2018,
  Smith2018}.  DSLs allow scientists their preferred approach of focusing on
science not software \citep{Kelly2007}.  A generative approach removes the
maintenance nightmare of documentation duplicates and near duplicates
\citep{LucivEtAl2018}, since knowledge is only captured once and automatically
transformed as needed.  Code generation has previously been applied to improve
SCS \citep{WhaleyEtAl2001, Veldhuizen1998, Pueschel2001}.  For instance, ATLAS
(Automatically Tuned Linear Algebra Software) \citep{WhaleyEtAl2001} and Blitz++
\citep{Veldhuizen1998} produces efficient and portable linear algebra software.
Spiral \citep{Pueschel2001} uses software/hardware generation for digital signal
processing.  \citet{Carette2008} shows how to generate a family of efficient,
type-safe Gaussian elimination algorithms.  FEniCS (Finite Element and
Computational Software) \citep{LoggEtAl2012} uses code generation when solving
differential equations.  Unlike previous work on SCS code generation, Drasil
(\href{https://github.com/JacquesCarette/Drasil}
{https://github.com/JacquesCarette/Drasil}) focus on generating all software
artifacts (requirements, design etc.), not just code.

Drasil, as presented in \citet{SzymczakEtAl2016} removes excuses for avoiding
documentation by providing transformative SCS development technology.  Drasil
provides an infrastructure for knowledge capture and document/code generation.
Drasil is implemented via Domain Specific Languages (DSLs) embedded in Haskell.
Drasil generates requirements documentation and code for several case studies,
including simulating the temperature of a solar water heating system, glass
breakage and two-dimensional game physics.  The full documentation
(requirements, design etc), code and test cases have been created manually for
each case study.  These manual case studies (many from
\citet{SmithJegatheesanAndKelly2016}) provide the `gold standard' against which
Drasil is tested.  All of these artifacts are available publicly.  The
generation techniques for Drasil began to take shape during work on generating
geometric data structures and functions \citep{CaretteEtAl2011}.

@@

roadmap - fit the different pieces together

@@

Key to the success of this exercise will be to involve and engage a partner for
each state of practice project.  In the previous effort we didn't involve domain
experts with the rationale of excluding potential bias, but this advantage is
not worth not being able to evaluate the functionality/usability of the software.
Moreover, not having an expert makes publication more difficult, since there is
no one to advice on how best to approach journals and publishers.  The domain
expert will be asked to help in the following ways:

@@

>>>>

DSLs and code/document generation provide a transformative technology for
documentation, design and verification \citep{JohansonAndHasselbring2018,
  Smith2018}.  Scientists prefer focusing on science not software
\citep{Kelly2007}.  DSLs allow them to do this.  A generative approach removes
the maintenance nightmare of documentation duplicates and near duplicates
\citep{LucivEtAl2018}, since knowledge is only captured once and automatically
transformed as needed.  Code generation has previously been applied to improve
SCS \citep{WhaleyEtAl2001, Veldhuizen1998, Pueschel2001}.  For instance, ATLAS
(Automatically Tuned Linear Algebra Software) \citep{WhaleyEtAl2001} and Blitz++
\citep{Veldhuizen1998} produces efficient and portable linear algebra software.
Spiral \citep{Pueschel2001} uses software/hardware generation for digital signal
processing.  \citet{Carette2008} shows how to generate a family of efficient,
type-safe Gaussian elimination algorithms.  FEniCS (Finite Element and
Computational Software) \citep{LoggEtAl2012} uses code generation when solving
differential equations.  Unlike previous work on SCS code generation, Drasil
(\href{https://github.com/JacquesCarette/Drasil}
{https://github.com/JacquesCarette/Drasil}) focus on generating all software
artifacts (requirements, design etc.), not just code.

Drasil, as presented in \citet{SzymczakEtAl2016} removes excuses for avoiding
documentation by providing transformative SCS development technology.  Drasil
provides an infrastructure for knowledge capture and document/code generation.
Drasil is implemented via Domain Specific Languages (DSLs) embedded in Haskell.
Drasil generates requirements documentation and code for several case studies,
including simulating the temperature of a solar water heating system, glass
breakage and two-dimensional game physics.  The full documentation
(requirements, design etc), code and test cases have been created manually for
each case study.  These manual case studies (many from
\citet{SmithJegatheesanAndKelly2016}) provide the `gold standard' against which
Drasil is tested.  All of these artifacts are available publicly at
\href{https://jacquescarette.github.io/Drasil/}
{https://jacquescarette.github.io/Drasil/}.  The generation techniques for
Drasil began to take shape during work on generating geometric data structures
and functions \citep{CaretteEtAl2011}.

@@

This section is decomposed into 4 subsections -- one for each of the different
approaches for assessing the impact of MDE on the sustainability of SCS.  The
goal of the first section is to establish a ``baseline'' on the current state of
the practice for SCS.  This is done by identifying domains within SCS where
significant software development work has been done and then systematically
assessing the quality of the existing work.  The next section looks specifically
at the impact of MDE on SCS through a series of case studies.  An ``ideal'' MDE
process is proposed and then followed with different stakeholders and
qualitative data is collected.  The last two sub-sections discuss assessing the
impact on SCS of assurance cases and computational variability testing,
respectively.

@@

Key to the success of this exercise will be to involve and engage a partner for
each state of practice project.  In the previous effort we did not involve domain
experts with the rationale of excluding potential bias, but this advantage is
not worth not being able to evaluate the functionality/usability of the software.
Moreover, not having an expert makes publication more difficult, since there is
no one to advice on how best to approach journals and publishers.  The domain
expert will be asked to help in the following ways:

@@

Write up experimental procedure as a publication - on arxiv, but hopefully
somewhere else too.

Start from previous work and modify.  Brainstorm how to do measures.

Schedule meeting.
